### 2018-07-10 source/Ch7-3C295/3c295.tex
backup

\chapter{Imaging 3C295 with International LOFAR}\label{section.3c295}
\minitoc
% % % % % % % % % % % % % % % % % % % % % % %
\section{Aims \& Methodology}
\pg
Our aim in this section is to create a high-resolution (0.1$''$) model of 3C295, which did not exist at LOFAR HBA frequencies when this project began. This means creating a model that will allow us to find good phase-calibration solutions for LOFAR international baselines, and thus use the international stations to create a high-resolution image of the EGS.
%While we also solve for amplitude gains, we know that we will need to correct the total source flux in each frequency subband as our initial spectral model is not necessarily correct.

@@

\pg
As outlined in \cref{section.RIME}, calibration is the process of solving for Jones matrices, which are associated to individual antennas. Here, the only source in our model (the source that dominates, by far, over other sources in the field) is 3C295; we can thus limit ourselves to \textit{direction-independent} calibration, which is equivalent to performing direction-dependent calibration in the direction of 3C295 only. The associated RIME can be written as:
\begin{align}
\Vis_{pq}^{t,\nu} &= \Gjones_p^{t,\nu} \left( \sum_{s=3C295} \Ejones_{sp}^{t,\nu} \Kjones_{sp}^{t,\nu} \Bmatrix \left(\Kjones_{sq}^{t,\nu}\right)^H \left(\Ejones_{sq}^{t,\nu}\right)^H \right) \left(\Gjones_q^{t,\nu}\right)^H\\
		  &= \Gjones_p^{t,\nu} \Ejones_{3C295,p}^{t,\nu} \Bmatrix_{3C295}^{t,\nu} \left(\Ejones_{3C295,q}^{t,\nu} \Gjones_q^{t,\nu}\right)^H\\
		  &= \Jones_{3C295,p}^{t,\nu} \Bmatrix_{3C295}^{t,\nu} \left(\Jones_{3C295,q}^{t,\nu}\right)^H\label{3c295.dirin.cal}
\end{align}
where we explicitly model the LOFAR beam during both calibration and imaging, so as to ensure that the true sensitivity of each baseline to this out-of-phase-centre calibrator is taken into account appropriately.

\pg
What interests us here is therefore the $\Jones_{3C295,p}^{t,\nu}$ for all international stations $p$ and at all times and frequencies $t,\nu$. We improve conditioning by using time and frequency solution intervals of a minute and 8 subbands; at those times where this interval is too large, the quality-based weighting scheme is expected to compensate. Choosing such large intervals also helps self-calibration converge faster. We do not explicitly resort to physics-based solutions to constrain the calibration solutions: we solve for the full 8 parameters of each complex $2\times 2$ Jones matrix.

\pg
We begin by using a simple spectral index of $\alpha=-0.8$, which is typical of synchrotron sources at these frequencies. It is not strictly correct, but gives a good first-order approximation of the integrated flux of 3C295 as a function of frequency. In so doing, we introduce a frequency- and baseline-dependent error factor:
\begin{align}
\hatJones_{3C295,p}^{t,\nu} &\approx \alpha_{pq}^\nu \Jones_{3C295,p}^{t,\nu}
\end{align}
where $\hatJones_{3C295,p}^{t,\nu}$ is our estimation of the Jones matrix $\Jones_{3C295,p}^{t,\nu}$ for time $t$ and frequency $\nu$, and the baseline-dependence of $\alpha_{pq}^\nu$ arises because 3C295 is not at phase centre. If it was, then $\alpha$ would be a smoothly-varying function of frequency alone.
Because we solve for the full Jones matrix, we cannot apply phase-only corrections\footnote{Indeed, what is the phase of a $2\times2$ matrix of complex numbers?} over multiple self-calibration rounds, as is standard practice. Over self-calibration rounds, we expect that $\alpha$ will change, but that it will remain smooth in frequency. The integrated flux of 3C295 over different passes of self-calibration is shown in \cref{fig.SC.ampfluct}: as this figure shows, the model amplitude can diverge significantly over the self-calibration process. %, sometimes even moving closer to the inital integrated flux, but often moving farther.

% images:
% IMAGES/3c295.fromVLA.6sb.withbeam.pass[1-5].int.model.fits IMAGES/3c295.fromVLA.60sb.withbeam.pass[1-3].int.model.fits IMAGES/MSlist.fullBW.LOBOS6.MSlist.fullBW.txt.uniform.int.model.fits
% script: /home/offler/plotSCamp.py
\begin{figure}[h]
\includegraphics[width=0.8\linewidth]{images/AmpFluct.png}
\caption{\label{fig.SC.ampfluct} Amplitude fluctuation of the models resulting from self-calibration iterations, expressed as a ratio of the model's integrated flux divided by the Scaife \& Healde integrated flux model at the appropriate frequency. Note that the integrated model flux for 324sb-pass1 (the penultimate data point in this plot) was lost to a careless command: the value shown here is just the average of the two neighbouring values.}
\end{figure}

\pg
Because $\alpha$ is smooth in frequency, we can apply a frequency-dependent amplitude correction (i.e. ``bootstrap") by relying on the work of Scaife \& Heald \citep[see][]{arse}. This correction will be done in the imaging plane, since $alpha$ is a function of $uv$-position (but not sky position). This is, again, because 3C295 does not lie at phase centre: if it did, the correction could be done equivalently in visibility or image space. By performing it in the image-plane, we rid ourselves of the $uv$-dependence. The resulting signal-to-noise improvement is shown in \cref{fig.SCrms}. It is calculated on the bootstrapped residual images resulting from each round of the self-calibration process - as such, the only thing that changes is the noise, since the signal is corrected to be the appropriate constant As we can see, the noise falls nicely with self-calibration passes. Tragically, the results from the first full-bandwidth calibration run with the 60-subband model were lost before this plot could be made, and so only the results from the second full-bandwidth self-calibration pass remain. We can see that it is a small improvement over the 60-subband model. 
\begin{figure}[h]
\includegraphics[width=0.8\linewidth]{images/SCrms.png}
\caption{\label{fig.SCrms} Bootstrapped residual RMS for each calibration pass. Note that the data for 324sb-pass1 (the penultimate data in this plot) were lost to a careless command: the value shown here is just the average of the two neighbouring values.}
\end{figure}

\pg
The large decrease (factor of $\sim2$)in RMS between the last 6-subband residual image and the first 60-subband residual image in \cref{fig.SCrms} is significant: we would expect a similar jump when going from 60 to 324 subbands. The fact that we do not is a strong indicator that we are digging into the limits of what our strategy can achieve by the end of the second full-bandwidth self-calibration pass. This is likely due to the fact that our initial sky model (and all subsequent models) only include 3C295: the fact that all the other sources in the primary beam are not included in the model is bound to introduce bias our gain solutions and artefacts in our images.

\pg
We want good phase and amplitude gain estimates without concern for the overall scaling factor: this means that we want to find the correct brightness distribution at a given frequency, knowing that the overall scaling factor will be wrong, and correcting the images accordingly afterwards. We self-calibrate over the full LOFAR bandwidth so that the spectral flux distribution (i.e. where the flux of 3C295 lies at different frequencies) can be recovered in greater depth as we introduce more and more of the full bandwidth. We do not begin self-calibration with the full bandwidth, as a single pass of self-calibration with 6 subbands takes about an hour and a half, whereas it takes a full 3 days with the full bandwidth. The full bandwidth is therefore only self-calibrated on once, at the very end. We begin our self-calibration with 6 subbands, chosen across the LOFAR bandwidth, calibrated simultaneously. The chosen subbands, and their position in the total bandwidth, are shown here:
\begin{figure}[h]
\begin{floatrow}
\ffigbox{%
  \includegraphics[width=0.5\textwidth]{images/FreqSampLabelled}
}{%
  \caption{\label{fig.freqsamp}Position of subbands chosen across full LOFAR bandwidth}%
}
\capbtabbox{%
  \begin{tabular}{|cc|} \hline
  Subband & $\nu_\mathrm{min} - \nu_\mathrm{max}$ \\ \hline
  SB035   &  117.45-117.69\\
  SB095   & 129.21-129.41\\
  SB155   & 140.93-141.12\\
  SB215   & 152.65-152.84\\
  SB275   & 164.37-164.56\\
  SB335   &  176.08-176.28 \\\hline
  \end{tabular}\vspace{1.2cm}
}{%
  \caption{Frequency bounds for the subbands chosen}%
}
\end{floatrow}
\end{figure}

@@

\pg
A good initial test of our hypothesis that we can start from a ``good enough" point to reach a global minimum is to self-calibrate on a two-point model. We perform this test, keeping only the two brightest pixels of the model shown above to test whether we converge to the same structure. The result is shown as an overlay in \cref{fig.sc.2pt}: as we can see, we recover the same structure as the full VLA model. We therefore start from this full model for our ``proper" self-calibration procedure.
\begin{figure}[h]
\includegraphics[width=0.8\linewidth]{images/{3C_295_I_8.7GHz_lbs2003.J2000.fits2ptmodel}.png}
\caption{\label{fig.sc.2pt} Overlay of the model resulting from a self-calibration process starting from a two-point model. The overlays rise exponentially, from the $5\sigma$ level to the maximum pixel value in the image. This image is very similar to the LOFAR image made with a single pass of calibration using the full VLA model.}
\end{figure}

@@

\pg
Let us begin by checking whether our core and remote station gains change over time due to our self-calibration procedure. Here, we look at the gains of SB155, on which we applied the initial and final gain solutions. This subband was included in the 60-subband self-calibration step, but not the 6-subband self-calibration step.
\begin{figure}[h!]
\begin{subfigure}{.9\textwidth}
\includegraphics[width=\linewidth]{images/{SB155.CS005HBA0.firstpass}.png}
\caption{\label{fig.sb155.gains.cs201.before} CS005, first calibration}
\end{subfigure}
\begin{subfigure}{.9\textwidth}
\includegraphics[width=\linewidth]{images/{SB155.CS005HBA0.lastpass}.png}
\caption{\label{fig.sb155.gains.cs201.after} CS005, last calibration}
\end{subfigure}
\begin{subfigure}{.9\textwidth}
\includegraphics[width=\linewidth]{images/{SB155.RS305HBA.firstpass}.png}
\caption{\label{fig.sb155.gains.RS305.before} RS305, first calibration}
\end{subfigure}
\begin{subfigure}{.9\textwidth}
\includegraphics[width=\linewidth]{images/{SB155.RS305HBA.lastpass}.png}
\caption{\label{fig.sb155.gains.RS305.afer} RS305, last calibration}
\end{subfigure}
\caption{\label{fig.sb155.gains.CR} Gain amplitude \& phase for CS201 of SB155}
\end{figure}

@@

\pg
As for the international station gains, they are shown below for three of the seven international stations used for our observation, in \cref{fig.sb155.gains.de602,fig.sb155.gains.se607,fig.sb155.gains.uk608}. We show only one of the four German stations. As we can see, while structure very clearly remains in some of these amplitude curves (notable in the last two hours, e.g. \cref{fig.sb155.gains.DE602.after}), some of the structure disappears from these amplitude curves. It should be noted that the last two hours of the observation are generally much noisier than the first 6; the last quarter of these plots show this. Quality-based weighting schemes compensate for this effect, and minimise the associated image deterioration. Note that the phase wraps extremely fast in the international station gain plots: this is because phase is calculated relative to CS001HBA0, the first antenna in our array. Because the international stations are much farther away from this core station than other core or remote stations are, we expect the associated phase solutions to wrap faster.
\begin{figure}[h!]
\begin{subfigure}{.9\linewidth}
\includegraphics[width=\linewidth]{images/{SB155.DE602HBA.firstpass}.png}
\caption{\label{fig.sb155.gains.DE602.before} DE602, first calibration}
\end{subfigure}
\begin{subfigure}{.9\linewidth}
\includegraphics[width=\linewidth]{images/{SB155.DE602HBA.lastpass}.png}
\caption{\label{fig.sb155.gains.DE602.after} DE602, last calibration}
\end{subfigure}
\caption{\label{fig.sb155.gains.de602} Gain amplitude \& phase for DE602, international stations of SB155}
\end{figure}

@@

\pg
So far, we have shown the gain solutions for one subband, out of a total of 324 used subbands. The spectral behaviour of the amplitude and phase solutions can give us insight into whether we are absorbing true flux or calibration artefacts into our model: if our calibration model becomes more accurate to the true underlying flux, we expect e.g. the phase of core station calibration solutions to remain the same (as we are improving high-resolution details of our model) while our international station calibration solutions would become flatter in amplitude. Similarly, we expect to see large-scale frequency-dependent behaviour in the phase solutions of international stations, with smaller fluctuations around this behaviour: the effect of differential TEC, ionospheric turbulence, etc. Indeed, while the core stations (and, to a lesser extent, the remote stations) can be approximated as having the ``same ionosphere" above them, this is not the case for international stations: this is a major source of the difficulties in calibrating these stations. As such, for international stations, we expect to see complex structure in the phase gain solutions, characteristic of the differences between the local atmosphere above the core and the international stations. For core stations, we only expect to see structure characteristic of scintillation in the phase gain solutions.

\pg
\cref{fig.fullBW.CS005,fig.fullBW.CS201HBA0} show the gain solution phases and amplitudes for all times and frequency in our observation. Some patches of the plot are empty (horizontal red bands): this is because some subbands had very poor data, and their use in self-calibration corrupted the final image. They are more numerous for the first calibration pass than the last, as the solutions were not sought for some cases.

\begin{figure}[h!]
\begin{subfigure}{.9\textwidth}
\includegraphics[width=\linewidth]{images/{CS005HBA0.firstpass.fullBW}.png}
\caption{\label{fig.fullBW.gains.CS005.before} CS005HBA0, first calibration, full bandwidth}
\end{subfigure}
\begin{subfigure}{.9\textwidth}
\includegraphics[width=\linewidth]{images/{CS005HBA0.lastpass.fullBW}.png}
\caption{\label{fig.fullBW.gains.CS005.after} CS005HBA0, last calibration, full bandwidth}
\end{subfigure}
\caption{\label{fig.fullBW.CS005} Plots showing the amplitude and phase solutions for CS005HBA0 at the start and end of self-calibration, respectively. Some areas are empty: in the case of the first calibration, this is because the solutions were not calculated for all subbands. For the last calibration, this is because some subbands were found to be corrupted, and their solutions are thus not shown here.}
\end{figure}

\begin{figure}[h!]
\begin{subfigure}{.9\textwidth}
\includegraphics[width=\linewidth]{images/{CS201HBA0.firstpass.fullBW}.png}
\caption{\label{fig.fullBW.CS201HBA0.before} CS201HBA0, first calibration, full bandwidth}
\end{subfigure}
\begin{subfigure}{.9\textwidth}
\includegraphics[width=\linewidth]{images/{CS005HBA0.lastpass.fullBW}.png}
\caption{\label{fig.fullBW.CS201HBA0.after} CS201HBA0, last calibration, full bandwidth}
\end{subfigure}
\caption{\label{fig.fullBW.CS201HBA0} Plots showing the amplitude and phase solutions for CS201HBA0 at the start and end of self-calibration, respectively. Some areas are empty: in the case of the first calibration, this is because the solutions were not calculated for all subbands. For the last calibration, this is because some subbands were found to be corrupted, and their solutions are thus not shown here.}
\end{figure}

\pg
As we can see, the behaviour of both phase and amplitude gains remains much the same for core stations, before and after self-calibration. This is a good sign: it means that we do not lose the large-scale structure in our observation as we self-calibrate. We see in \cref{fig.fullBW.RS305HBA} that remote stations have much more phase structure as a function of frequency. This structure shows that the ionosphere above the remote Dutch stations is different from that above the core stations: this is an expected result. The fact that the structure does not change significantly before and after the self-calibration run indicates that the remote station gains, like the core stations gains, do not absorb true flux into their gain solutions during the self-calibration process. It can thus be taken as a sign that our self-calibration is trustworthy for these stations.
\begin{figure}[h!]
\begin{subfigure}{.9\textwidth}
\includegraphics[width=\linewidth]{images/{RS305HBA.firstpass.fullBW}.png}
\caption{\label{fig.fullBW.RS305HBA.before} RS305HBA, first calibration, full bandwidth}
\end{subfigure}
\begin{subfigure}{.9\textwidth}
\includegraphics[width=\linewidth]{images/{RS305HBA.lastpass.fullBW}.png}
\caption{\label{fig.fullBW.RS305HBA.after} RS305HBA, last calibration, full bandwidth}
\end{subfigure}
\caption{\label{fig.fullBW.RS305HBA} Plots showing the amplitude and phase solutions for RS305HBA at the start and end of self-calibration, respectively. Some areas are empty: in the case of the first calibration, this is because the solutions were not calculated for all subbands. For the last calibration, this is because some subbands were found to be corrupted, and their solutions are thus not shown here.}
\end{figure}

@@

\begin{figure}[h!]
\begin{subfigure}{.9\textwidth}
\includegraphics[width=\linewidth]{images/{DE601HBA.firstpass.fullBW}.png}
\caption{\label{fig.fullBW.DE601HBA.before} DE601HBA, first calibration, full bandwidth}
\end{subfigure}
\begin{subfigure}{.9\textwidth}
\includegraphics[width=\linewidth]{images/{DE601HBA.lastpass.fullBW}.png}
\caption{\label{fig.fullBW.DE601HBA.after} DE601HBA, last calibration, full bandwidth}
\end{subfigure}
\caption{\label{fig.fullBW.DE601HBA} Plots showing the amplitude and phase solutions for DE601HBA at the start and end of self-calibration, respectively. Some areas are empty: in the case of the first calibration, this is because the solutions were not calculated for all subbands. For the last calibration, this is because some subbands were found to be corrupted, and their solutions are thus not shown here.}
\end{figure}

>>>>

\chapter{Imaging 3C295 with International LOFAR}\label{section.3c295}
\minitoc
% % % % % % % % % % % % % % % % % % % % % % %
\section{Aims \& Methodology}
\pg
Our aim in this section is to create a high-resolution (0.1$''$) model of 3C295, which did not exist at LOFAR HBA frequencies when this project began. This means creating a model that will allow us to find good phase-calibration solutions for LOFAR international baselines, and thus use the international stations to create a high-resolution image of the EGS. Because we only solve for gains in the direction of 3C295, we must then estimate the impact of directional gain errors over the rest of the field.
%While we also solve for amplitude gains, we know that we will need to correct the total source flux in each frequency subband as our initial spectral model is not necessarily correct.

@@

\subsection{Our RIME}

\pg
As outlined in \cref{section.RIME}, calibration is the process of solving for Jones matrices, which are associated to individual antennas. Here, the only source in our model (the source that dominates, by far, over other sources in the field) is 3C295; we can thus limit ourselves to \textit{direction-independent} calibration, which is equivalent to performing direction-dependent calibration in the direction of 3C295 only. The associated RIME can be written as:
\begin{align}
\Vis_{pq}^{\nu} &= \Gjones_p^{\nu} \left( \sum_{s=3C295} \Ejones_{sp}^{\nu} \Kjones_{sp}^{\nu} \Bmatrix \left(\Kjones_{sq}^{\nu}\right)^H \left(\Ejones_{sq}^{\nu}\right)^H \right) \left(\Gjones_q^{\nu}\right)^H\\
		  &= \Gjones_p^{\nu} \Ejones_{3C295,p}^{\nu} \Bmatrix_{3C295}^{\nu} \left(\Ejones_{3C295,q}^{\nu} \Gjones_q^{\nu}\right)^H\\
		  &= \Jones_{3C295,p}^{\nu} \Bmatrix_{3C295}^{\nu} \left(\Jones_{3C295,q}^{\nu}\right)^H\label{3c295.dirin.cal}
\end{align}
where we mute the time-dependence of all terms, and explicitly model the LOFAR beam during both calibration and imaging, so as to ensure that the true sensitivity of each baseline to this out-of-phase-centre calibrator is taken into account appropriately.

\pg
What interests us here is therefore the $\Jones_{3C295,p}^{\nu}$ for all international stations $p$ and at all times and frequencies $t,\nu$. We improve conditioning by using time and frequency solution intervals of a minute and 8 subbands; at those times where this interval is too large, the quality-based weighting scheme is expected to compensate. Choosing such large intervals also helps self-calibration converge faster. We do not explicitly resort to physics-based solutions to constrain the calibration solutions: we solve for the full 8 parameters of each complex $2\times 2$ Jones matrix.

\subsection{Calibration Strategy}

\pg
We self-calibrate over the full LOFAR bandwidth so that the spectral flux distribution (i.e. where the flux of 3C295 lies at different frequencies) can be recovered in greater depth as we introduce more and more of the full bandwidth. We do not begin self-calibration with the full bandwidth, as a single pass of self-calibration with 6 subbands takes about an hour and a half, whereas it takes a full 3 days with the full bandwidth. The full bandwidth is therefore only self-calibrated on once, at the very end. We begin our self-calibration with 6 subbands, chosen across the LOFAR bandwidth, calibrated simultaneously. The chosen subbands, and their position in the total bandwidth, are shown here:
\begin{figure}[h]
\begin{floatrow}
\ffigbox{%
  \includegraphics[width=0.5\textwidth]{images/FreqSampLabelled}
}{%
  \caption{\label{fig.freqsamp}Position of subbands chosen across full LOFAR bandwidth}%
}
\capbtabbox{%
  \begin{tabular}{|cc|} \hline
  Subband & $\nu_\mathrm{min} - \nu_\mathrm{max}$ \\ \hline
  SB035   &  117.45-117.69\\
  SB095   & 129.21-129.41\\
  SB155   & 140.93-141.12\\
  SB215   & 152.65-152.84\\
  SB275   & 164.37-164.56\\
  SB335   &  176.08-176.28 \\\hline
  \end{tabular}\vspace{1.2cm}
}{%
  \caption{Frequency bounds for the subbands chosen}%
}
\end{floatrow}
\end{figure}

@@

\subsection{Testing the impact of initial model choice}

\pg
A good initial test of our hypothesis that we can start from a ``good enough" point to reach a global minimum is to self-calibrate on a two-point model. We perform this test, keeping only the two brightest pixels of the model shown above to test whether we converge to the same structure. The result is shown as an overlay in \cref{fig.sc.2pt}: as we can see, we recover the same structure as the full VLA model. We appear to have introduced a small astrometric error, however: this is likely an indicator that using only the peak flux, rather than the full distribution, will bias the ``centre of flux" away from its true position -- hence introducing an astrometric shift. We therefore start from the full model for our ``proper" self-calibration procedure.
\begin{figure}[h]
\includegraphics[width=0.8\linewidth]{images/{3C_295_I_8.7GHz_lbs2003.J2000.fits2ptmodel}.png}
\caption{\label{fig.sc.2pt} Overlay of the model resulting from a self-calibration process starting from a two-point model. The overlays rise exponentially, from the $5\sigma$ level to the maximum pixel value in the image. This image is very similar to the LOFAR image made with a single pass of calibration using the full VLA model.}
\end{figure}

@@

\subsection{Frequency-dependent Scale Error}

\pg
We begin by using a simple spectral index of $\alpha=-0.8$, which is typical of synchrotron sources at these frequencies. It is not strictly correct, but gives a good first-order approximation of the integrated flux of 3C295 as a function of frequency, which is shown in \cref{fig.3c295.intflux}
\begin{figure}[h]
\includegraphics[width=0.8\linewidth]{images/3C295-intflux.png}
\caption{\label{fig.3c295.intflux} Integrated flux of 3C295 as a function of frequency, as fitted by \citepads{arse}.}
\end{figure}

In so doing, we introduce a frequency-dependent error factor in our gain solutions:
\begin{align}
\hatJones_{3C295,p}^{\nu} &\approx \sqrt{\alpha_\nu} \Jones_{3C295,p}^{\nu}
\end{align}
where $\hatJones_{3C295,p}^{\nu}$ is our estimation of the Jones matrix $\Jones_{3C295,p}^{\nu}$ for time $t$ and frequency $\nu$. $\alpha_\nu$ is a smoothly-varying function of frequency alone. This then introduces a frequency-dependent error factor $1/\sqrt{\alpha_\nu}$ in our visibilities and therefore in our images. This factor is a constant for all visibilities at a given frequency.

\pg
Because we solve for the full Jones matrix, we cannot apply phase-only corrections\footnote{Indeed, what is the phase of a $2\times2$ matrix of complex numbers?} over multiple self-calibration rounds, as is standard practice. Over self-calibration rounds, we expect that $\alpha$ will change, but that it will remain smooth in frequency. The integrated flux of 3C295 over different passes of self-calibration is shown in \cref{fig.SC.ampfluct}: as this figure shows, the model amplitude can diverge significantly over the self-calibration process. %, sometimes even moving closer to the inital integrated flux, but often moving farther.

% images:
% IMAGES/3c295.fromVLA.6sb.withbeam.pass[1-5].int.model.fits IMAGES/3c295.fromVLA.60sb.withbeam.pass[1-3].int.model.fits IMAGES/MSlist.fullBW.LOBOS6.MSlist.fullBW.txt.uniform.int.model.fits
% script: /home/offler/plotSCamp.py
\begin{figure}[h!]
\includegraphics[width=0.8\linewidth]{images/AmpFluct.png}
\caption{\label{fig.SC.ampfluct} Amplitude fluctuation of the models resulting from self-calibration iterations, expressed as a ratio of the model's integrated flux divided by the Scaife \& Healde integrated flux model at the appropriate frequency. Note that the integrated model flux for 324sb-pass1 (the penultimate data point in this plot) was lost to a careless command: the value shown here is just the average of the two neighbouring values.}
\end{figure}

\pg
Because $\alpha$ is smooth in frequency, we can apply a frequency-dependent amplitude correction (i.e. ``bootstrap") by relying on the work of Scaife \& Heald \citep[see][]{arse}. This correction will be done in the imaging plane, since $alpha$ is a function of $uv$-position (but not sky position). This is, again, because 3C295 does not lie at phase centre: if it did, the correction could be done equivalently in visibility or image space. By performing it in the image-plane, we rid ourselves of the $uv$-dependence. The resulting signal-to-noise improvement is shown in \cref{fig.SCrms}. It is calculated on the bootstrapped residual images resulting from each round of the self-calibration process - as such, the only thing that changes is the noise, since the signal is corrected to be the appropriate constant As we can see, the noise falls nicely with self-calibration passes. Tragically, the results from the first full-bandwidth calibration run with the 60-subband model were lost before this plot could be made, and so only the results from the second full-bandwidth self-calibration pass remain. We can see that it is a small improvement over the 60-subband model. 
\begin{figure}[h!]
\includegraphics[width=0.8\linewidth]{images/SCrms.png}
\caption{\label{fig.SCrms} Bootstrapped residual RMS for each calibration pass. Note that the data for 324sb-pass1 (the penultimate data in this plot) were lost to a careless command: the value shown here is just the average of the two neighbouring values.}
\end{figure}

\pg
The large decrease (factor of $\sim2$)in RMS between the last 6-subband residual image and the first 60-subband residual image in \cref{fig.SCrms} is significant: we would expect a similar jump when going from 60 to 324 subbands. The fact that we do not is a strong indicator that we are digging into the limits of what our strategy can achieve by the end of the second full-bandwidth self-calibration pass: we are stuck in a local minimum. This is likely due to the fact that our initial sky model (and all subsequent models) only include 3C295: the fact that all the other sources in the primary beam are not included in the model is bound to introduce bias our gain solutions and artefacts in our images.

\pg
We want good phase and amplitude gain estimates without concern for the overall scaling factor: this means that we want to find the correct brightness distribution at a given frequency, knowing that the overall scaling factor will be wrong, and correcting the images accordingly afterwards.

\subsection{Examining gain solutions}

\pg
Let us begin by checking whether our core and remote station gains change over time due to our self-calibration procedure. Here, we look at the gains of SB155, on which we solved for gain solutions using our initial and final models. This subband was included in the 60-subband self-calibration step, but not the 6-subband self-calibration step.
\begin{figure}[h!]
\begin{subfigure}{.9\textwidth}
\includegraphics[width=\linewidth]{images/{SB155.CS005HBA0.firstpass}.png}
\caption{\label{fig.sb155.gains.cs201.before} CS005, first calibration}
\end{subfigure}
\begin{subfigure}{.9\textwidth}
\includegraphics[width=\linewidth]{images/{SB155.CS005HBA0.lastpass}.png}
\caption{\label{fig.sb155.gains.cs201.after} CS005, last calibration}
\end{subfigure}
\begin{subfigure}{.9\textwidth}
\includegraphics[width=\linewidth]{images/{SB155.RS305HBA.firstpass}.png}
\caption{\label{fig.sb155.gains.RS305.before} RS305, first calibration}
\end{subfigure}
\begin{subfigure}{.9\textwidth}
\includegraphics[width=\linewidth]{images/{SB155.RS305HBA.lastpass}.png}
\caption{\label{fig.sb155.gains.RS305.afer} RS305, last calibration}
\end{subfigure}
\caption{\label{fig.sb155.gains.CR} Gain amplitude \& phase for CS201 of SB155}
\end{figure}

@@

\pg
As for the international station gains, they are shown below for three of the seven international stations used for our observation, in \cref{fig.sb155.gains.de602,fig.sb155.gains.se607,fig.sb155.gains.uk608}. We show only one of the four German stations. As we can see, while structure very clearly remains in some of these amplitude curves (notable in the last two hours, e.g. \cref{fig.sb155.gains.DE602.after}), some of the structure disappears from these amplitude curves. It should be noted that the last two hours of the observation are generally much noisier than the first 6; the last quarter of these plots show this. Quality-based weighting schemes compensate for this effect, and minimise the associated image deterioration. Note that the phase wraps extremely fast in the international station gain plots: this is because phase is calculated relative to CS001HBA0, the first antenna in our array. Because the international stations are much farther away from this core station than other core or remote stations are, we expect the associated phase solutions to wrap faster.
\begin{figure}[h!]
\begin{subfigure}{.9\linewidth}
\includegraphics[width=\linewidth]{images/{SB155.DE602HBA.firstpass}.png}
\caption{\label{fig.sb155.gains.DE602.before} DE602, first calibration}
\end{subfigure}
\begin{subfigure}{.9\linewidth}
\includegraphics[width=\linewidth]{images/{SB155.DE602HBA.lastpass}.png}
\caption{\label{fig.sb155.gains.DE602.after} DE602, last calibration}
\end{subfigure}
\caption{\label{fig.sb155.gains.de602} Gain amplitude \& phase for DE602 of SB155}
\end{figure}

@@

\pg
So far, we have shown the gain solutions for one subband, out of a total of 324 used subbands. The spectral behaviour of the amplitude and phase solutions can give us insight into whether we are absorbing true flux or calibration artefacts into our model: if our calibration model becomes more accurate to the true underlying flux, we expect e.g. the phase of core station calibration solutions to remain the same (as we are improving high-resolution details of our model) while our international station calibration solutions would become flatter in amplitude. Similarly, we expect to see large-scale frequency-dependent behaviour in the phase solutions of international stations, with smaller fluctuations around this behaviour: the effect of differential TEC, ionospheric turbulence, etc. Indeed, while the core stations (and, to a lesser extent, the remote stations) can be approximated as having the ``same ionosphere" above them, this is not the case for international stations: this is a major source of the difficulties in calibrating these stations. As such, for international stations, we expect to see complex structure in the phase gain solutions, characteristic of the differences between the local atmosphere above the core and the international stations. For core stations, we only expect to see structure characteristic of scintillation (regular stripes with no frequency-dependent structure) in the phase gain solutions.

\pg
\cref{fig.fullBW.CS005,fig.fullBW.CS201HBA0} show the gain solution phases and amplitudes for all times and frequencies in our observation. Some patches of the plot are empty (horizontal red bands): this is because some subbands had very poor data, and their use in self-calibration corrupted the final image. They are more numerous for the first calibration pass than the last, as the solutions were not sought for some cases.

\begin{figure}[h!]
\begin{subfigure}{.9\textwidth}
\includegraphics[width=\linewidth]{images/{CS005HBA0.firstpass.fullBW}.png}
\caption{\label{fig.fullBW.gains.CS005.before} CS005HBA0, first calibration, full bandwidth}
\end{subfigure}
\begin{subfigure}{.9\textwidth}
\includegraphics[width=\linewidth]{images/{CS005HBA0.lastpass.fullBW}.png}
\caption{\label{fig.fullBW.gains.CS005.after} CS005HBA0, last calibration, full bandwidth}
\end{subfigure}
\caption{\label{fig.fullBW.CS005} Plots showing the amplitude and phase solutions for CS005HBA0 at the start and end of self-calibration, respectively. The black areas correspond to regions which were not used for that pass of calibration due to poor data.}
\end{figure}

\begin{figure}[h!]
\begin{subfigure}{.9\textwidth}
\includegraphics[width=\linewidth]{images/{CS201HBA0.firstpass.fullBW}.png}
\caption{\label{fig.fullBW.CS201HBA0.before} CS201HBA0, first calibration, full bandwidth}
\end{subfigure}
\begin{subfigure}{.9\textwidth}
\includegraphics[width=\linewidth]{images/{CS005HBA0.lastpass.fullBW}.png}
\caption{\label{fig.fullBW.CS201HBA0.after} CS201HBA0, last calibration, full bandwidth}
\end{subfigure}
\caption{\label{fig.fullBW.CS201HBA0} Plots showing the amplitude and phase solutions for CS201HBA0 at the start and end of self-calibration, respectively. The black areas correspond to regions which were not used for that pass of calibration due to poor data.}
\end{figure}

\pg
As we can see, the behaviour of both phase and amplitude gains remains much the same for core stations, before and after self-calibration. This is a good sign: it means that we do not lose the large-scale structure in our observation as we self-calibrate. We see in \cref{fig.fullBW.RS305HBA} that remote stations have much more phase structure as a function of frequency. This structure shows that the ionosphere above the remote Dutch stations is different from that above the core stations: this is an expected result. The fact that the structure does not change significantly before and after the self-calibration run could indicate that the remote station gains, like the core stations gains, do not absorb true flux into their gain solutions during the self-calibration process. It can thus be taken as a sign that our self-calibration is trustworthy for these stations.
\begin{figure}[h!]
\begin{subfigure}{.9\textwidth}
\includegraphics[width=\linewidth]{images/{RS305HBA.firstpass.fullBW}.png}
\caption{\label{fig.fullBW.RS305HBA.before} RS305HBA, first calibration, full bandwidth}
\end{subfigure}
\begin{subfigure}{.9\textwidth}
\includegraphics[width=\linewidth]{images/{RS305HBA.lastpass.fullBW}.png}
\caption{\label{fig.fullBW.RS305HBA.after} RS305HBA, last calibration, full bandwidth}
\end{subfigure}
\caption{\label{fig.fullBW.RS305HBA} Plots showing the amplitude and phase solutions for RS305HBA at the start and end of self-calibration, respectively. The black areas correspond to regions which were not used for that pass of calibration due to poor data.}
\end{figure}

@@

\begin{figure}[h!]
\begin{subfigure}{.9\textwidth}
\includegraphics[width=\linewidth]{images/{DE601HBA.firstpass.fullBW}.png}
\caption{\label{fig.fullBW.DE601HBA.before} DE601HBA, first calibration, full bandwidth}
\end{subfigure}
\begin{subfigure}{.9\textwidth}
\includegraphics[width=\linewidth]{images/{DE601HBA.lastpass.fullBW}.png}
\caption{\label{fig.fullBW.DE601HBA.after} DE601HBA, last calibration, full bandwidth}
\end{subfigure}
\caption{\label{fig.fullBW.DE601HBA} Plots showing the amplitude and phase solutions for DE601HBA at the start and end of self-calibration, respectively. The black areas correspond to regions which were not used for that pass of calibration due to poor data.}
\end{figure}

\begin{figure}[h!]
\begin{subfigure}{.9\textwidth}
\includegraphics[width=\linewidth]{images/{DE602HBA.firstpass.fullBW}.png}
\caption{\label{fig.fullBW.DE602HBA.before} DE602HBA, first calibration, full bandwidth}
\end{subfigure}
\begin{subfigure}{.9\textwidth}
\includegraphics[width=\linewidth]{images/{DE602HBA.lastpass.fullBW}.png}
\caption{\label{fig.fullBW.DE602HBA.after} DE602HBA, last calibration, full bandwidth}
\end{subfigure}
\caption{\label{fig.fullBW.DE602HBA} Plots showing the amplitude and phase solutions for DE602HBA at the start and end of self-calibration, respectively. The black areas correspond to regions which were not used for that pass of calibration due to poor data.}
\end{figure}
